But I can't preregister my research: Improving the reproducibility and transparency of ecology and conservation with adaptive preregistration for model-based research

### 2025-03-03 _analysis/ms.qmd
rm superfluous bullet in abstract

  3.	We developed a generalised ‘modelling workflow’ to capture critical analysis decisions in the model development cycle, informing a preregistration template for model-based research. The iterative nature of model development is incompatible with existing, linear preregistration templates, creating a barrier to implementing preregistration for model-based research.
  4.	To better align the internal logic of preregistration with the iterative and non-linear process of ecological modelling, we propose, test and evaluate a methodology for ‘adaptive preregistration’, using a case-study of environmental flows modelling for decision-support in Victoria, Australia.
  5. This research provides a template and methodology for implementing adaptive preregistration of ecological models. Although we focus on ecology and conservation in this paper, the concept of adaptive preregistration, and the templates developed here, could be applied in model-based research in other scientific disciplines within science more broadly. Modellers need no longer cry “but I can’t preregister my research.”

@@

| Preregistration Version | Preregistration Process | Modelling and Analysis Process |
|----|----|----|
| **Version 1:** (v0.10 – v0.8.1) | Establish problem context, study background and aims, describe data collection and data cleaning process, articulate modelling objectives, and specify candidate models. | Data cleaning and preparation undertaken concurrently by lead and supporting researchers due to limited resourcing and short timelines for project deliverables. Candidate models addressing different components of the research question were preregistered. Identify data structural properties and potential consequences for model fitting |
| **Version 2:** (v0.9.0-v.0.12.5) | Describe Pilot Analysis on data subset and specify registered flexibility in the form of decision heuristics. | Critical uncertainties in modelling decisions were articulated (e.g. how do we classify best flow regime based on inundation data?). In addition, data partitioning, exploratory data analysis, refined candidate models and model checking procedure preregistered |
| **Version 3:** (v0.13.0-v.0.20.1) | Describe Main Analysis, including specification of decision tree for model selection process based on results of pilot analysis. | Conducted modelling and analysis as outlined in preregistration. Reported deviations and rationale as necessary. |
|  |  |  |

>>>>

  3.	To better align the internal logic of preregistration with the iterative and non-linear process of ecological modelling, we propose, test and evaluate a methodology for ‘adaptive preregistration’, using a case-study of environmental flows modelling for decision-support in Victoria, Australia.
  4. This research provides a template and methodology for implementing adaptive preregistration of ecological models. Although we focus on ecology and conservation in this paper, the concept of adaptive preregistration, and the templates developed here, could be applied in model-based research in other scientific disciplines within science more broadly. Modellers need no longer cry “but I can’t preregister my research.”

@@

| Preregistration Version | Preregistration Process | Modelling and Analysis Process |
|------------------------|------------------------|------------------------|
| **Version 1:** (v0.10 – v0.8.1) | Establish problem context, study background and aims, describe data collection and data cleaning process, articulate modelling objectives, and specify candidate models. | Data cleaning and preparation undertaken concurrently by lead and supporting researchers due to limited resourcing and short timelines for project deliverables. Candidate models addressing different components of the research question were preregistered. Identify data structural properties and potential consequences for model fitting |
| **Version 2:** (v0.9.0-v.0.12.5) | Describe Pilot Analysis on data subset and specify registered flexibility in the form of decision heuristics. | Critical uncertainties in modelling decisions were articulated (e.g. how do we classify best flow regime based on inundation data?). In addition, data partitioning, exploratory data analysis, refined candidate models and model checking procedure preregistered |
| **Version 3:** (v0.13.0-v.0.20.1) | Describe Main Analysis, including specification of decision tree for model selection process based on results of pilot analysis. | Conducted modelling and analysis as outlined in preregistration. Reported deviations and rationale as necessary. |
|  |  |  |
